---
title: "Mazda3 Analysis"
author: "Daley Mikalson"
date: "August 1, 2016"
output:  
      html_document:  
        keep_md: true 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The data

For some reason, I kept all the records related to my car for all the years that I owned it. Now that I have sold the car, I am going to clear all the paper out, but not before seeing what it tells me about what the car cost me.

I used a fuel tracker for iOS called [MileageKeeper](http://mileagekeeper.com "Mileage Keeper") and it worked well enough that I didn't ever look for a new one. It has the ability to export and import .csv files, which was quite helpful when Apple had to give me a new phone with a functioning camera.

I have all the service records at a couple different garages in Ottawa and Toronto. This includes oil changes, swapping winter/summer tires and repairs.

```{r fuel}
fuel <- read.csv("data/Mazda3.csv")
summary(fuel)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
