---
title: "Mazda3 Analysis"
author: "Daley Mikalson"
date: "August 2, 2016"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      dpi=250)
library(dplyr)
library(magrittr)
library(ggplot2)
library(lubridate)
```

## The data
For some reason, I kept all the records related to my car for all the years that I owned it. Now that I have sold the car, I am going to clear all the paper out, but not before seeing what it tells me about what the car cost me.

I used a fuel tracker for iOS called [MileageKeeper](http://mileagekeeper.com "Mileage Keeper") and it worked well enough that I didn't ever look for a new one. It has the ability to export and import .csv files, which was quite helpful when Apple had to give me a new phone with a functioning camera.

I have all the service records at a couple different garages in Ottawa and Toronto. This includes oil changes, swapping winter/summer tires and repairs. The few times that I had to pay for parking are also in my saved data.

The only thing that is missing is all the random times that I stopped for car washes.

```{r data}
fuel <- read.csv("data/Mazda3.csv")

fuel$Distance <- c(-diff(fuel$Odometer),NA)
fuel$L.100km <- fuel$Fuel.Amount / fuel$Distance * 100
summary(fuel)

service <- read.csv("data/Service.csv")
summary(service)
fixed <- read.csv("data/Fixed.csv")
summary(fixed)
parking <- read.csv("data/Parking.csv")
summary(parking)

```

Let's look at the total for each cost area.

```{r summarise}
totals <- summarise(group_by(fixed, Type), Cost = sum(Cost))
totals <- rbind(totals, summarise(group_by(service, Type), Cost = sum(Cost)))
totals <- rbind(totals, data.frame(Type = "Fuel", Cost = sum(fuel$Total.Cost)))
totals <- rbind(totals, data.frame(Type = "Parking", Cost = sum(parking$Cost)))
totals
sum(totals$Cost)
```

```{r bars, echo = FALSE}
p <- ggplot(totals, aes(x = reorder(Type, -Cost), y = Cost))
p + geom_bar(stat = "identity", position = "identity") +
  xlab("Type of Cost") + theme(axis.text.x = element_text(angle=90))
```

Rather than looking at things by type of cost, we could look at it by year. It is useful to also break it out by Fixed vs. Variable. The really big bar in 2014 is mostly explained by me paying off the balance of the car so that I would stop paying 4.99% interest each year.
```{r annual, echo = FALSE}
a <- summarise(group_by(fixed, year(Date)), Cost = sum(Cost), Kind = "fixed")
b <- summarise(group_by(service, year(Date)), Cost = sum(Cost), Kind = "variable")
c <- summarise(group_by(fuel, year(Date)), Cost = sum(Total.Cost), Kind = "variable")
d <- summarise(group_by(parking, year(Date)), Cost = sum(Cost), Kind = "variable")

t <- bind_rows(list(a,b,c,d))
names(t)[1] = "Year"

p <- ggplot()
p + geom_bar(data = subset(t, Cost >= 0), aes(x = Year, y= Cost, fill = Kind), stat= "identity") + 
  geom_bar(data = subset(t, Cost < 0), aes(x = Year, y= Cost, fill = Kind), stat= "identity", position = "identity") +
  scale_y_continuous(breaks = seq(-3000, 11000, 1000))
```


You can also embed plots, for example:

```{r pressure, echo=FALSE}
q <- ggplot(fuel, aes(x=Odometer, y=Fuel.Amount))
q + geom_bar(stat="identity")
```
